---
title: "Untitled"
author: "NYEW JIA WEN`"
date: "2025-03-12"
output: html_document
---

```{r}
setwd("~/ST3189 ML/cousework")
```

##Regression tasks

```{r cars}
library(glmnet)
library(rpart)
library(randomForest)
library(caTools)# for train-test split
library(leaps)#for backward elimination
library(corrplot)
library(MASS)
library(car)#for vif（）to check multicollinearity
```

DATASET LINK: [link](https://archive.ics.uci.edu/dataset/275/bike+sharing+dataset)
just take day.csv as dataset
[link](https://www.kaggle.com/datasets/vjchoudhary7/customer-segmentation-tutorial-in-python?resource=download)

```{r}
# Load RQ1 Dataset (day file for daily rentals)
day <- read.csv("day.csv", stringsAsFactors = TRUE)
summary(day)

```

```{r}
# Converting relevant columns to factors for RQ1
day$season <- as.factor(day$season) 
day$yr <- as.factor(day$yr)          # yr: 0: 2011, 1: 2012
day$mnth <- as.factor(day$mnth)      # mnth: 1: January, 2: February, ..., 12: December
day$holiday <- as.factor(day$holiday) # holiday: 0: No, 1: Yes
day$weekday <- as.factor(day$weekday) 
day$workingday <- as.factor(day$workingday) # workingday: 0: No, 1: Yes
day$weathersit <- as.factor(day$weathersit) # weathersit: 1: Clear, 2: Few clouds, 3: Cloudy, 4: Rainy

# Check the data structure
str(day)

```


##LINEAR REGRESSION MODEL FOR REGRESSION TASK

```{r}
# EDA part for RQ1-DAY 

nuMeric.day <- day[, sapply(day, is.numeric)] 

correlation <- cor(nuMeric.day, use = "complete.obs")

corrplot(correlation, method = "color", type = "upper", tl.col = "black", tl.srt = 45, addCoef.col = "grey")


```

##aim to find Backward Elimination RMSE ##RQ1

```{r}
# Train-Test Split (70% train, 30% test)
set.seed(2025)
split <- sample.split(day$cnt, SplitRatio = 0.7) 
trainset <- subset(day, split == T)
testset <- subset(day, split == F)

```

```{r}
bikecountlinear <- lm(cnt ~ temp + atemp + hum + windspeed + season + weathersit + holiday + workingday, data = trainset)
summary(bikecountlinear)

```

```{r}
# Backward Elimination using step from MASS package
#backward elimination to remove the least significant predictors step by step.
BE <- step(bikecountlinear, direction = "backward")
summary(BE)
options(digits=3)
coef(BE)

```

```{r}
RMSE.BE.training <- sqrt(mean(residuals(BE)^2))
RMSE.BE.training

predict.BE.test <- predict(BE, newdata = testset)
testset.error <- testset$cnt - predict.BE.test
RMSE.BE.test <- sqrt(mean((testset.error)^2))
print(paste("RMSE for Backward Elimination Test Set:", RMSE.BE.test))

```

##RQ1 10-fold CV with linear regression

```{r}
# Load necessary libraries
library(caret)

# Ensure train-test split and folds are created properly
set.seed(2025)  # Set seed for reproducibility

# Create 10 folds using the caret package for better split handling
folds <- createFolds(day$cnt, k = 10, list = TRUE, returnTrain = TRUE)


RMSE.train <- numeric(10)
RMSE.test <- numeric(10)

# 10FCV
for (i in 1:10) {
  trainset <- day[folds[[i]], ]  # Training data
  testset <- day[-folds[[i]], ]  # Test data

  bikecountlinear <- lm(cnt ~ temp + atemp + hum + windspeed + season + weathersit + holiday + workingday, data = trainset)

  # RMSE-TRAINSET
  train_pred <- predict(bikecountlinear, newdata = trainset)
  RMSE.train[i] <- sqrt(mean((trainset$cnt - train_pred)^2))

  # RMSE-TESTSET
  test_pred <- predict(bikecountlinear, newdata = testset)
  RMSE.test[i] <- sqrt(mean((testset$cnt - test_pred)^2))
}

# Compute the mean RMSE for both train and test sets
avg.rmse.training <- mean(RMSE.train)
avg.rmse.testing <- mean(RMSE.test)

# Output the RMSE values
print(paste("Mean RMSE for Training Set:", avg.rmse.training))
print(paste("Mean RMSE for Testing Set:", avg.rmse.testing))

# Calculate the standard error of RMSE for both train and test sets
RMSE.train.se <- sd(RMSE.train) / sqrt(10)
RMSE.test.se <- sd(RMSE.test) / sqrt(10)

RMSE.train.se; RMSE.test.se

```

```{r}
par(mfrow=c(2,2))
plot(bikecountlinear)
par(mfrow = c(1,1))

```

```{r}
vif(BE)
#No major multicollinearity issues are apparent since the VIF values are generally below 5

```

##Ridge and Lasso Regression ##RQ1 

```{r}
X <- day[, !(names(day) %in% c("cnt"))]  # All columns except target 'cnt'
y <- day$cnt

# Train-Test Split (70% train, 30% test)
set.seed(2025)
split <- sample.split(y, SplitRatio = 0.7)
trainset <- subset(day, split == T)
testset <- subset(day, split == F)

# Prepare X for Ridge and Lasso models (standardize X)
# Ensure all columns in X are numeric
X.train <- as.data.frame(trainset[, !(names(trainset) %in% c("cnt"))])
X.test <- as.data.frame(testset[, !(names(testset) %in% c("cnt"))])

# Convert factors to numeric using model.matrix(because just now error for matrix problem)
train.numeric.mtrx <- model.matrix(~., data = X.train)[, -1] 
test.numeric.mtrx<- model.matrix(~., data = X.test)[, -1]  

train.dt.mtrx <- as.matrix(train.numeric.mtrx)
test.dt.mtrx<- as.matrix(test.numeric.mtrx)


```
# Ridge Regression -------------------------------------
```{r}
grid <- 10^seq(10, -2, length = 100)
ridge.model <- glmnet(train.dt.mtrx, trainset$cnt, alpha = 0, lambda = grid)
cv.ridge <- cv.glmnet(train.dt.mtrx, trainset$cnt, alpha = 0)
plot(cv.ridge)

```

```{r}
bl.ridge <- cv.ridge$lambda.min
ridge.predictions <- predict(ridge.model, s = bl.ridge, newx = test.dt.mtrx)
RMSE.ridge <- sqrt(mean((ridge.predictions - testset$cnt)^2))
print(paste("RMSE for Ridge Regression:", RMSE.ridge))
```
# Lasso Regression
```{r}
lA.model <- glmnet(train.dt.mtrx, trainset$cnt, alpha = 1, lambda = grid)
cv.lA <- cv.glmnet(train.dt.mtrx, trainset$cnt, alpha = 1)
plot(cv.lA)

```

```{r}
bl.lA <- cv.lA$lambda.min
lA.predictions <- predict(lA.model, s = bl.lA, newx = test.dt.mtrx)
RMSE.lA <- sqrt(mean((lA.predictions - testset$cnt)^2))
print(paste("RMSE for Lasso Regression:", RMSE.lA))
```

##10-fold Cross Validation for Ridge and Lasso Regression 

```{r}
# 10-Fold Cross Validation (Train-Test Split inside CV)
set.seed(2025)  
folds <- createFolds(y, k = 10, list = T, returnTrain = T)


RMSE.ridge.train <- numeric(10)
RMSE.ridge.test <- numeric(10)
RMSE.lA.train <- numeric(10)
RMSE.lA.test <- numeric(10)

# 10-Fold CV loop
for (i in 1:10) {
  
  trainset <- day[folds[[i]], ]  # Training data
  testset <- day[-folds[[i]], ]  # Test data
  
  # Prepare X for Ridge and Lasso models (standardize X)
  X.train <- as.data.frame(trainset[, !(names(trainset) %in% c("cnt"))])
  X.test <- as.data.frame(testset[, !(names(testset) %in% c("cnt"))])

  # Convert factors to numeric using model.matrix
  train.numeric.mtrx <- model.matrix(~., data = X.train)[, -1] 
  test.numeric.mtrx <- model.matrix(~., data = X.test)[, -1]  

  train.dt.mtrx <- as.matrix(train.numeric.mtrx)
  test.dt.mtrx <- as.matrix(test.numeric.mtrx)

  cv.ridge <- cv.glmnet(train.dt.mtrx, trainset$cnt, alpha = 0)  
  bl.ridge <- cv.ridge$lambda.min 
  ridge.predictions.train <- predict(cv.ridge, s = bl.ridge, newx = train.dt.mtrx)
  RMSE.ridge.train[i] <- sqrt(mean((ridge.predictions.train - trainset$cnt)^2))  # RMSE for training set
  
  ridge.predictions.test <- predict(cv.ridge, s = bl.ridge, newx = test.dt.mtrx)
  RMSE.ridge.test[i] <- sqrt(mean((ridge.predictions.test - testset$cnt)^2))  # RMSE-test set
  
  # Lasso Regression (alpha = 1 for Lasso)
  cv.lA<- cv.glmnet(train.dt.mtrx, trainset$cnt, alpha = 1)  
  bl.lA <- cv.lA$lambda.min  
  lA.predictions.train <- predict(cv.lA, s = bl.lA, newx = train.dt.mtrx)
  RMSE.lA.train[i] <- sqrt(mean((lA.predictions.train - trainset$cnt)^2))  # RMSE for training set
  
  lA.predictions.test <- predict(cv.lA, s = bl.lA, newx = test.dt.mtrx)
  RMSE.lA.test[i] <- sqrt(mean((lA.predictions.test - testset$cnt)^2))  # RMSE for test set
}

```

```{r}
# Compute the mean RMSE for both Ridge and Lasso models across all 10 folds
RMSE.ridge.train.mean <- mean(RMSE.ridge.train)
RMSE.ridge.test.mean <- mean(RMSE.ridge.test)
RMSE.lA.train.mean <- mean(RMSE.lA.train)
RMSE.lA.test.mean <- mean(RMSE.lA.test)

# Output the results
cat("Mean RMSE for Ridge Regression (Trainset):", RMSE.ridge.train.mean, "\n")
cat("Mean RMSE for Ridge Regression (Testset):", RMSE.ridge.test.mean, "\n")
cat("Mean RMSE for Lasso Regression (Trainset):", RMSE.lA.train.mean, "\n")
cat("Mean RMSE for Lasso Regression (Testset):", RMSE.lA.test.mean, "\n")

```

```{r}
#Ridge
bl.ridge <- cv.ridge$lambda.min
print(paste("Best Lambda for Ridge Regression:", bl.ridge))

# Lasso
bl.lA <- cv.lA$lambda.min
print(paste("Best Lambda for Lasso Regression:", bl.lA))

```

###CART MODEL-REGRESSION TASK
```{r}

library(rpart)
library(rpart.plot)


set.seed(2025)

# prePrepare (using same predictors as Linear Regression)
X <- day[, c("temp", "atemp", "hum", "windspeed", "season", "weathersit", "holiday", "workingday")]  # Select predictors
y <- day$cnt  # The target variable (total daily bike rentals)

# Train-Test Split (70% train, 30% test)
set.seed(2025)
split <- sample.split(y, SplitRatio = 0.7)
trainset <- subset(day, split == T)
testset <- subset(day, split == F)

```

```{r}
# Fit CART for regression using selected predictors
bikecountcart<- rpart(cnt ~ temp + atemp + hum + windspeed + season + weathersit + holiday + workingday, 
                    data = trainset, 
                    method = "anova", 
                    control = rpart.control(minsplit = 2, cp = 0))

```

```{r}
# Print CP table to identify pruning criteria
printcp(bikecountcart)
```
```{r}
cp.opt <- bikecountcart$cptable[which.min(bikecountcart$cptable[, "xerror"]), "CP"]
cp.opt
```

```{r}
#choose optimal cp value from printcp
cp.opt <- 0.00601  
cart.pruned <- prune(bikecountcart, cp = cp.opt)


rpart.plot(cart.pruned, nn = T, main = "Pruned CART Model")


cart.predictions <- predict(cart.pruned, newdata = testset)

```

```{r}

RMSE.cart <- sqrt(mean((cart.predictions - testset$cnt)^2))
print(paste("RMSE for CART Regression (Test Set):", RMSE.cart))

```

```{r}
library(caret)
library(rpart)

set.seed(2025)
folds <- createFolds(day$cnt, k = 10)

RMSE.cart.train <- numeric(10)
RMSE.cart.test <- numeric(10)

for (i in 1:10) {
  trainset <- day[folds[[i]], ]
  testset <- day[-folds[[i]], ]
  
  # Train full tree
  bikecountcart<- rpart(cnt ~ temp + atemp + hum + windspeed + season + weathersit + holiday + workingday,
                      data = trainset,
                      method = "anova",
                      control = rpart.control(minsplit = 2, cp = 0))
  
  cp.opt <- bikecountcart$cptable[which.min(bikecountcart$cptable[, "xerror"]), "CP"]
  cart.pruned <- prune(bikecountcart, cp = cp.opt)
  
  # Predictions
  pred.train <- predict(cart.pruned, newdata = trainset)
  pred.test <- predict(cart.pruned, newdata = testset)
  
  # RMSE
  RMSE.cart.train[i] <- sqrt(mean((trainset$cnt - pred.train)^2))
  RMSE.cart.test[i] <- sqrt(mean((testset$cnt - pred.test)^2))
}

# Average RMSE
mean(RMSE.cart.train)
mean(RMSE.cart.test)

```


##RANDOM FOREST MODEL FOR REGRESSION TASK

```{r}
##Error: The number of rows in X.train.matrix does not match the length of the response variable(settled)
# Load necessary libraries
library(randomForest)

# Prepare the data (same as used in other models)
X <- day[, c("temp", "atemp", "hum", "windspeed", "season", "weathersit", "holiday", "workingday")]  # Select predictors
y <- day$cnt  # The target variable (total daily bike rentals)

# Train-Test Split (70% train, 30% test)
set.seed(2025)
split <- sample.split(y, SplitRatio = 0.7)
trainset <- subset(day, split == T)
testset <- subset(day, split == F)

# Ensure training and testing sets are in the correct format (no need for model.matrix transformation here for Random Forest)
X.train <- trainset[, c("temp", "atemp", "hum", "windspeed", "season", "weathersit", "holiday", "workingday")]
X.test <- testset[, c("temp", "atemp", "hum", "windspeed", "season", "weathersit", "holiday", "workingday")]


bikecountrf<- randomForest(x = X.train, y = trainset$cnt, ntree = 500, importance = TRUE)

rf.pre <- predict(bikecountrf, newdata = X.test)

# RMSE-TESTSET
RMSE.rf <- sqrt(mean((rf.pre - testset$cnt)^2))
print(paste("RMSE for Random Forest Regression (Test Set):", RMSE.rf))

print("Feature Importance for Random Forest:")
print(importance(bikecountrf))

```

```{r}
# Display the feature importance
var.impt <- importance(bikecountrf)
View(var.impt)
varImpPlot(bikecountrf,type=1)
```
## CLASSIFICATION TASKS (LOGISTIC REGRESSION, KNN, SVM,NN) - SUPERVISED LEARNING
DATASET LINK: [link](https://www.kaggle.com/c/titanic/data)
just take train as dataset

```{r}
# Load required libraries
library(caret)
library(pROC)
library(ggplot2)
library(nnet)
library(e1071)
library(class)
library(data.table)
library(caTools)
library(reshape2)
# Load dataset
tit <- fread("train.csv")
summary(tit)
#age has lots na
```

##Logistic Regression
```{r}

library(data.table)
library(caTools)
library(pROC)
library(ggplot2)
library(reshape2)

# Function to compute classification metrics
compute.metrics <- function(cm) {
  TruePo <- cm[2,2]#实际是 1，预测也是 1 python要加1
  FalsePo <- cm[1,2]#实际是 0，但预测成 1
  FalseNeg <- cm[2,1]#实际是 1，但预测成 0
  TrueNeg <- cm[1,1]#实际是 0，预测也是 0
  
  accuracy <- (TruePo + TrueNeg) / sum(cm)
  precision <- TruePo / (TruePo + FalsePo)
  recall <- TruePo / (TruePo + FalseNeg)
  F1 <- 2 * (precision * recall) / (precision + recall)
  
  return(list(Accuracy = accuracy, Precision = precision, Recall = recall, F1.Score = F1))
}


```

```{r}
tit$Survived <- factor(tit$Survived)
tit$Pclass <- factor(tit$Pclass)
tit$Sex <- factor(tit$Sex, levels = c("male", "female"))
tit$Embarked <- factor(tit$Embarked)


tit$Age[is.na(tit$Age)] <- median(tit$Age, na.rm = T)
tit$Fare[is.na(tit$Fare)] <- median(tit$Fare, na.rm = T)
tit <- na.omit(tit)#所以这一步可能会删掉大量资料。危险
#最好先只保留这些变量，再处理缺失。


set.seed(2025)
train <- sample.split(Y = tit$Survived, SplitRatio = 0.7)
trainset <- subset(tit, train == T)
testset <- subset(tit, train == F)

# Logistic Regression 
logistic<- glm(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, 
                      family = binomial, data = trainset)
summary(logistic)

# Compute Odds Ratios
#因为 logistic regression 的系数原本是在 log-odds 尺度上，不直观。
#所以你把它 exponentiate，变成 odds ratio。
OR.logistic <- exp(coef(logistic))
#OR > 1：这个变量增加，生还 odds 上升
#OR < 1：这个变量增加，生还 odds 下降
OR.CI.logistic <- exp(confint(logistic))#把置信区间也转成 odds ratio 的尺度，方便解释

# Predictions
prob.test <- predict(logistic, newdata = testset, type = 'response')
#type = "response" 会给你 预测概率，范围在 0 到 1。
predict.test <- ifelse(prob.test > 0.5, 1, 0)

# Convert predictions to factor
predict.test <- factor(predict.test, levels = c(0,1))#方便后面做 confusion matrix。

# Confusion Matrix
cm <- table(testset$Survived, predict.test)#建立混淆矩阵：行 = 实际值，列 = 预测值
print(cm)

# Compute Metrics Using Function
metrics <- compute.metrics(cm)
print(metrics)

##Error :cannot use logistic regression
```

##KNN
```{r}
# Convert categorical variables to numeric for KNN
#因为 KNN 不能直接处理 factor/categorical variable，所以你先把类别变量转成 numeric。
#更好的做法：用 dummy variables / one-hot encoding 会更好
tit$Sex <- as.numeric(tit$Sex)
tit$Embarked <- as.numeric(tit$Embarked)
tit$Pclass <- as.numeric(tit$Pclass)


set.seed(2025)
train <- sample.split(Y = tit$Survived, SplitRatio = 0.7)
trainset <- subset(tit, train == TRUE)
testset <- subset(tit, train == FALSE)


train.X <- trainset[, c("Pclass", "Sex", "Age", "SibSp", "Parch", "Fare", "Embarked"), with = FALSE]
test.X <- testset[, c("Pclass", "Sex", "Age", "SibSp", "Parch", "Fare", "Embarked"), with = FALSE]
train.Y <- trainset$Survived
test.Y <- testset$Survived
#这里要注意什么
#KNN 非常依赖距离，所以通常要先 standardize / normalize。
#你这里没有做 scaling，这会让像 Fare、Age 这种数值范围大的变量主导距离计算。

# Run KNN model
set.seed(2025)
k <- 5 #这是一个常见起点，后续可以通过交叉验证调参，不然老师/面试官可能会问：为什么是 5，不是 3 或 7？
knn.pred <- knn(train.X, test.X, train.Y, k = k)


cm <- table(test.Y, knn.pred)
print(cm)


metrics <- compute.metrics(cm)
print(metrics)


roc.tt <- roc(as.numeric(test.Y), as.numeric(knn.pred))#problem：KNN 的 ROC 最好用“概率”而不是硬分类结果
plot(roc.test, col = "blue", main = "ROC Curve - KNN")#problem：这里应该选roc.tt的
auc.tt <- auc(roc.tt)
cat("AUC:", auc.tt, "\n")


```

```{r}
# Heatmap CM
#melt：把 table 转成长格式 dataframe
cm.melt <- melt(cm)
colnames(cm.melt) <- c("True Label", "Predicted Label", "Count")
ggplot(cm.melt, aes(x = `Predicted Label`, y = `True Label`, fill = Count)) + 
  geom_tile() + #geom_tile()->画方块热图
  geom_text(aes(label = Count), color = "white", size = 5) + #geom_text(aes(label = Count)):把数量写在格子里
  scale_fill_gradient(low = "lightblue", high = "darkblue") + 
  labs(title = "Confusion Matrix - KNN", x = "Predicted labels", y = "True labels") + 
  theme_minimal()


```

```{r}
cat("\nK-Nearest Neighbors (KNN)\n")
cat("F1 Score:", round(metrics$F1.Score, 3), "\n")
cat("ROC_AUC Score:", round(auc.test, 3), "\n")#这一行应该会报错：auc.tt
cat("Accuracy:", round(metrics$Accuracy, 3), "\n")
```

##SVM
```{r}
# Convert categorical variables to numeric for SVM
#跟 KNN 一样，把分类变量先转成数值，好让 SVM 可以使用。
#as.numeric  还是会有“类别被强行赋顺序”的问题。通常 dummy encoding 会更稳。
tit$Sex <- as.numeric(tit$Sex)
tit$Embarked <- as.numeric(tit$Embarked)
tit$Pclass <- as.numeric(tit$Pclass)


set.seed(2025)
train <- sample.split(Y = tit$Survived, SplitRatio = 0.7)
trainset <-subset(tit, train == T)
testset <- subset(tit, train == F)

# Define training and test data
train.X <- trainset[, c("Pclass", "Sex", "Age", "SibSp", "Parch", "Fare", "Embarked"), with = FALSE]
test.X <- testset[, c("Pclass", "Sex", "Age", "SibSp", "Parch", "Fare", "Embarked"), with = FALSE]
train.Y <- trainset$Survived
test.Y <- testset$Survived

# SVM Model
set.seed(2025)
svmm<- svm(x = train.X, y = train.Y, kernel = "radial", gamma = 1, cost = 1, probability = T)
#gamma 大，边界更复杂；gamma 小，边界更平滑。
#radial:表示这个分类边界不是简单直线，而是可以做非线性分类
#cost 大，模型更想把训练集分对；但可能更容易 overfit
summary(svmm)

# Hyperparameter tuning
set.seed(2025)
train.svm <- cbind(train.X, Survived = train.Y)

tune.out <- tune(svm, Survived ~ ., data = train.svm, kernel = "radial",
                 ranges = list(cost = c(0.1, 1, 10, 100, 1000), gamma = c(0.5, 1, 2, 3, 4)))
#让他去测很多
summary(tune.out)


best.svm <- tune.out$best.model#保存最佳模型

#后面预测时并没有用 best.svm，而是还在用原本的 svmm。这就有点浪费了调参步骤。
# Predictions
svm.pred <- predict(svmm, test.X, probability = T)#predict(best.svm, test.X, probability = TRUE)
predicted.probs <- attr(svm.pred, "probabilities")[,2]#从预测结果里把“属于 class 1”的概率取出来。
predict.class <- ifelse(predicted.probs > 0.5, 1, 0)


predict.class <- factor(predict.class, levels = c(0,1))


```

```{r}
# Confusion Matrix#solve error
if (length(test.Y) == length(predict.class)) {
  cm <- table(test.Y, predict.class)
  print(cm)
} else {
  cat("Error: Mismatched lengths - test.Y:", length(test.Y), " predict.class:", length(predict.class), "\n")
}

print(cm)#这样还是可能报错。所以这段逻辑不够完整。

# Compute Metrics Using Function
metrics <- compute.metrics(cm)
print(metrics)


```

```{r}
# ROC Curve and AUC
#把真实标签转 numeric，然后算 ROC / AUC。
# Ensure test.Y is numeric and check lengths
test.Y.numeric <- as.numeric(as.character(test.Y))

if (length(test.Y.numeric) == length(predicted.probs) && all(!is.na(predicted.probs))) {
  roc.tt <- roc(test.Y.numeric, predicted.probs)
  plot(roc.test, col = "blue", main = "ROC Curve - SVM")#应该是roc.tt
  auc.tt <- auc(roc.tt)
  cat("AUC:", auc.tt, "\n")
} else {
  cat("Error: Mismatched lengths or missing values - test.Y:", length(test.Y.numeric), 
      " predicted.probs:", length(predicted.probs), "\n")
}

plot(roc.tt, col = "blue", main = "ROC Curve - SVM")
auc.tt <- auc(roc.tt)
cat("AUC:", auc.tt, "\n")


```

```{r}
# Heatmap cm
cm.melt <- melt(cm)
colnames(cm.melt) <- c("True Label", "Predicted Label", "Count")
ggplot(cm.melt, aes(x = `Predicted Label`, y = `True Label`, fill = Count)) + 
  geom_tile() + 
  geom_text(aes(label = Count), color = "white", size = 5) + 
  scale_fill_gradient(low = "lightblue", high = "darkblue") + 
  labs(title = "Confusion Matrix - SVM", x = "Predicted labels", y = "True labels") + 
  theme_minimal()


```

```{r}
cat("\nSupport Vector Machine (SVM)\n")
cat("F1 Score:", round(metrics$F1.Score, 3), "\n")
cat("ROC_AUC Score:", round(auc.test, 3), "\n")#这里应该要roc.tt
cat("Accuracy:", round(metrics$Accuracy, 3), "\n")
```

##NN
```{r}
library(data.table)
library(neuralnet)
library(pROC)

#Data preprocessing
tit$Age[is.na(tit$Age)] <- median(tit$Age, na.rm = TRUE)
tit$Fare[is.na(tit$Fare)] <- median(tit$Fare, na.rm = TRUE)
tit <- tit[!is.na(Embarked), ]#比前面好 也更合理

# select feature
tit <- tit[, .(Survived, Pclass, Sex, Age, SibSp, Parch, Fare, Embarked)]#只保留神经网络要用的变量，避免无关列干扰。

#先把类别变量设成 factor，方便后面做 dummy variable。
tit$Survived <- as.factor(tit$Survived)
tit$Pclass <- as.factor(tit$Pclass)
tit$Sex <- as.factor(tit$Sex)
tit$Embarked <- as.factor(tit$Embarked)

# Create dummy variables
tit$Pclass1 <- ifelse(tit$Pclass == 1, 1, 0)#是否为一等舱
tit$Pclass2 <- ifelse(tit$Pclass == 2, 1, 0)#没有单独建 Pclass3，因为它自动当 baseline
tit$SexMale <- ifelse(tit$Sex == "male", 1, 0)#没有单独建 female，因为 female 当 baseline
tit$EmbarkedC <- ifelse(tit$Embarked == "C", 1, 0)#是否从 C 港口上船
tit$EmbarkedQ <- ifelse(tit$Embarked == "Q", 1, 0)#没有单独建 S，表示 S 当 baseline

# Normalize numerical 
#把数值变量做 min-max normalization，压到 0 到 1 之间。
#因为：神经网络对变量尺度比较敏感。如果不缩放，某些数值大的变量会影响训练。
tit$Age <- (tit$Age - min(tit$Age)) / (max(tit$Age) - min(tit$Age))
tit$Fare <- (tit$Fare - min(tit$Fare)) / (max(tit$Fare) - min(tit$Fare))
tit$SibSp <- (tit$SibSp - min(tit$SibSp)) / (max(tit$SibSp) - min(tit$SibSp))
tit$Parch <- (tit$Parch - min(tit$Parch)) / (max(tit$Parch) - min(tit$Parch))

# Prepare input data
input_data <- tit[, .(Age, Fare, SibSp, Parch, Pclass1, Pclass2, SexMale, EmbarkedC, EmbarkedQ)]
target <- tit$Survived

# Convert target variable to numeric
target <- as.numeric(as.character(target))

```

```{r}
set.seed(2025)
nn <- neuralnet(
  target ~ Age + Fare + SibSp + Parch + Pclass1 + Pclass2 + SexMale + EmbarkedC + EmbarkedQ,
  data = input_data,#有bug
  hidden = 2,  # 1个隐藏层，2个节点
  linear.output = FALSE,#因为这是分类问题，不是回归，所以输出层不是线性的
  err.fct = "ce",#因为这是分类问题，不是回归，所以输出层不是线性的
  act.fct = "logistic"#因为这是分类问题，不是回归，所以输出层不是线性的
)

#但是公式左边的 target 不在 input_data 里。这通常会报错，因为模型找不到 target 这列。
#应该nn_data <- cbind(input_data, target = target)
#data = nn_data
#你这里是直接拿 整个 input_data 来预测。
#也就是说，你没有做 NN 专属的 train/test split。
#这样会导致：你是在训练资料上预测自己结果会偏乐观,不能公平跟前面的 logistic / KNN / SVM 比
# Plot neural network structure
plot(nn)


```

```{r}
# predict
prob <- compute(nn, input_data)
predicted <- ifelse(prob$net.result > 0.5, 1, 0)


cm <- table(target, predicted)
print(cm)

# Calculate performance metrics
TruePo <- cm[2, 2]
FalsePo <- cm[1, 2]
FalseNeg <- cm[2, 1]
TrueNeg <- cm[1, 1]

accuracy <- (TruePo + TrueNeg) / sum(cm)
precision <- TruePo / (TruePo + FalsePo)
recall <- TruePo / (TruePo + FalseNeg)
F1 <- 2 * (precision * recall) / (precision + recall)

# Output
cat("Neural Network Performance:\n")
cat("Accuracy:", round(accuracy, 3), "\n")
cat("Precision:", round(precision, 3), "\n")
cat("Recall:", round(recall, 3), "\n")
cat("F1 Score:", round(F1, 3), "\n")

```

```{r}
# ROC curve (with fixes)
prob_vector <- prob$net.result[, 1]  # Extract numeric vector from matrix
rocobjective <- roc(target, prob_vector, direction = "<")  # Set direction explicitly
plot(rocobjective, col = "blue", main = "ROC Curve")
aucval <- auc(rocobjective)
cat("AUC:", round(aucval, 3), "\n")
```
##EDA-classification
```{r}
# Barplot of survival
ggplot(tit, aes(x = Survived)) +
  geom_bar(aes(fill = Survived), width = 0.7) +
  labs(title = "Survival Count", x = "Survived", y = "Count") +
  scale_fill_manual(values = c("red", "blue")) +
  theme_minimal()

```

```{r}
# Pclass and survival relationship
ggplot(tit, aes(x = Pclass, fill = Survived)) +
  geom_bar(position = "fill", width = 0.7) +
  labs(title = "Survival Rate by Pclass", x = "Pclass", y = "Survival Rate") +
  scale_fill_manual(values = c("red", "blue")) +
  theme_minimal()

```

```{r}

ggplot(tit, aes(x = Age, y = Fare, color = Sex)) +
  geom_point(alpha = 0.6) +
  labs(title = "Age vs Fare by Sex", x = "Age", y = "Fare") +
  theme_minimal()

```


## UNSUPERVISED LEARNING TASKS (PCA, K-MEANS CLUSTERING, HIERARCHICAL CLUSTERING)
Dataset Link:
[link](https://www.kaggle.com/datasets/vjchoudhary7/customer-segmentation-tutorial-in-python?resource=download)
```{r}
df <- read.csv("Mall_Customers.csv")
head(df)

```

```{r}
summary(df)
```
##EDA 
```{r}
library(ggplot2)
library(corrplot)
library(GGally)
library(gridExtra)



# Data manipulation 
# Converting 'Gender' column to numeric (Male=0, Female=1)
df$Gender <- ifelse(df$Gender == "Male", 0, 1)

# Scaling 
df.scaling <- df[, c("Age", "Annual.Income..k..", "Spending.Score..1.100.")]
df.scaling <- scale(df.scaling)

# Handling missing values 
for(i in 1:ncol(df.scaling)){
  df.scaling[is.na(df.scaling[,i]), i] <- mean(df.scaling[,i], na.rm = T)
}

sum(is.na(df.scaling))  

# Assign the 'CustomerID' column as the row names for better data handling
rownames(df.scaling) <- df$CustomerID

# Remove the 'CustomerID' column as it's not needed for clustering
df.scaling <- as.data.frame(df.scaling)

```

```{r}
# **1. Correlation Matrix with Heatmap**
cor_matrix <- cor(df.scaling)
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.cex = 0.8, tl.col = "black", title = "Correlation Matrix", 
         addCoef.col = "white", number.cex = 0.8)


```

```{r}
# **2. Boxplots for Distribution Visualization**
p1 <- ggplot(df, aes(x = "", y = Age)) + 
      geom_boxplot(fill = "skyblue", color = "black") + 
      ggtitle("Boxplot: Age") + 
      theme_minimal()

p2 <- ggplot(df, aes(x = "", y = Annual.Income..k..)) + 
      geom_boxplot(fill = "skyblue", color = "black") + 
      ggtitle("Boxplot: Annual Income (k$)") + 
      theme_minimal()

p3 <- ggplot(df, aes(x = "", y = Spending.Score..1.100.)) + 
      geom_boxplot(fill = "skyblue", color = "black") + 
      ggtitle("Boxplot: Spending Score (1-100)") + 
      theme_minimal()

# **Displaying the Boxplots in a Grid**
grid.arrange(p1, p2, p3, ncol = 3, nrow = 1)


```

```{r}
# **3. Pairplot (Scatterplot Matrix)**
pairplot <- ggpairs(df.scaling, 
                    title = "Pairplot of Numerical Features", 
                    aes(color = factor(df$Gender), alpha = 0.5))

# Print the Pairplot (separate print, as ggpairs is a grid object)
print(pairplot)
```

##PCA
```{r}

out <- prcomp(df.scaling, scale = T)
out$rotation

out$x


out$sdev  
pr.var <- out$sdev^2  
pve <- pr.var / sum(pr.var)  
```

```{r}
# Load necessary library
library(ggplot2)

# Standardize the data 
df.scaling <- scale(df[, c("Age", "Annual.Income..k..", "Spending.Score..1.100.")])


out <- prcomp(df.scaling, scale = T)


pr.var <- out$sdev^2

pve <- pr.var / sum(pr.var)

# Create Scree Plot
ggplot(data = data.frame(PC = 1:length(pve), Variance = pve), aes(x = PC, y = Variance)) +
  geom_point() +
  geom_line() +
  xlab("Principal Component") +
  ylab("Proportion of Variance Explained") +
  ggtitle("Scree Plot: PCA") +
  theme_minimal()

```

```{r}
# Print the standard deviations and calculate the proportion of variance explained
pr.var <- out$sdev^2
pve <- pr.var / sum(pr.var)


pve

```


```{r}
# **PCA Biplot**
biplot(out, scale = 0, cex = 0.7, col = "black", 
       xlab = "PC1", ylab = "PC2")
# Add arrows and labels only for the most significant variables prevent messy
text(out$rotation[,1], out$rotation[,2], labels = rownames(out$rotation), 
     col = "red", cex = 0.6, pos = 3)
```

##KNN
```{r}
library(ggplot2)
library(cluster)
library(dplyr)

#Perform K-Means Clustering with different numbers of clusters (1 to 9)
inertia <- numeric(length = 9)  # To store inertia (sum of squared errors)
K <- 1:9

# Set the seed before running K-means
set.seed(2025)  

# Loop to calculate inertia for each value of K (1 to 9)
for (i in K) {
  kmeans.model <- kmeans(df.scaling, centers = i, nstart = 25)  # nstart ensures good initialization
  inertia[i] <- kmeans.model$tot.withinss
}

```

```{r}
# Plot the Elbow Method (Scree Plot)
plot(K, inertia, type = "b", pch = 19, col = "blue", 
     main = "Scree Plot: Elbow Method", xlab = "Number of Clusters", ylab = "Sum of Squared Error", 
     cex.lab = 1.2, cex.main = 1.5)
abline(v = 3, col = "red", lty = 2)  # The elbow at K=3 is marked with a red dashed line

```

```{r}
# Fit  (K=3)
kmeans.model <- kmeans(df.scaling, centers = 3, nstart = 25)  # K=3 based on elbow method


kmeans.df <- as.data.frame(df)
kmeans.df$kclusters <- as.factor(kmeans.model$cluster)  # Add cluster labels as a new column

# Preview the top rows of the data with cluster labels
head(kmeans.df)


```

```{r}
# Plotting Boxplots for each variable against the clusters
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))

# Boxplot for Age vs K-Means clusters
boxplot(kmeans.df$Age ~ kmeans.df$kclusters, 
        main = "Age vs kclusters", 
        xlab = "K-Clusters", ylab = "Age", 
        col = "lightblue", border = "black")

# Boxplot for Annual Income (k$) vs K-Means clusters
boxplot(kmeans.df$Annual.Income..k.. ~ kmeans.df$kclusters, 
        main = "Annual Income (k$) vs kclusters", 
        xlab = "K-Clusters", ylab = "Annual Income (k$)", 
        col = "lightgreen", border = "black")

# Boxplot for Spending Score vs K-Means clusters
boxplot(kmeans.df$Spending.Score..1.100. ~ kmeans.df$kclusters, 
        main = "Spending Score vs kclusters", 
        xlab = "K-Clusters", ylab = "Spending Score", 
        col = "lightpink", border = "black")

# Display the plots
par(mfrow = c(1, 1))  


```

```{r}
#summary
kmeans.table <- kmeans.df %>% group_by(kclusters) %>% summarise_all(list(mean = ~mean(.)))
print(kmeans.table)
```

##HAC
```{r}

library(ggplot2)
library(cluster)
library(dendextend)

# Perform hierarchical clustering using different linkage methods
comp <- hclust(dist(df.scaling), method = "complete")
avg <- hclust(dist(df.scaling), method = "average")
sin<- hclust(dist(df.scaling), method = "single")
cen <- hclust(dist(df.scaling), method = "centroid")
ward <- hclust(dist(df.scaling), method = "ward.D2")

# Improved Dendrogram  with Coloring and Cutting at k=3
par(mfrow = c(3, 2))  # Adjust the layout for multiple plots

# Plot dendrograms with better labels and color coding for clusters
plot(comp, main = "Complete Linkage", xlab = "", sub = "", cex = 0.8)
rect.hclust(comp, k = 3, border = 2:4) 

plot(avg, main = "Average Linkage", xlab = "", sub = "", cex = 0.8)
rect.hclust(avg, k = 3, border = 2:4)  

plot(sin, main = "Single Linkage", xlab = "", sub = "", cex = 0.8)
rect.hclust(sin, k = 3, border = 2:4)  

plot(cen, main = "Centroid Linkage", xlab = "", sub = "", cex = 0.8)
rect.hclust(cen, k = 3, border = 2:4)  

plot(ward, main = "Ward Linkage", xlab = "", sub = "", cex = 0.8)
rect.hclust(ward, k = 3, border = 2:4)  
par(mfrow = c(1, 1))  


```

```{r}
# Calculate Agglomerative Coefficients
ac_comp <- agnes(df.scaling, method = "complete")$ac
ac_avg <- agnes(df.scaling, method = "average")$ac
ac_sin <- agnes(df.scaling, method = "single")$ac
ac_ward <- agnes(df.scaling, method = "ward")$ac

ac_coef_df <- data.frame(
  single = ac_sin, 
  complete = ac_comp, 
  average = ac_avg, 
  ward = ac_ward
)

print(ac_coef_df)

# Ward method has the highest agglomerative coefficient, so we'll use Ward for cutting the tree.

```

```{r}
# Dendrogram for Ward method with enhanced coloring and labels
ward.d <- as.dendrogram(ward)  # Convert hclust to dendrogram for better control
ward.d <- color_branches(ward.d, k = 3)  
ward.d <- color_labels(ward.d, k = 3)  
plot(ward.d, main = "Ward Linkage with Enhanced Coloring", cex = 0.8)

```

```{r}
# Cutting the dendrogram based on the number of clusters (k=7 and k=8)
sil <- cutree(ward, k = 7)  # Using k=7 from Silhouette method
gap <- cutree(ward, k = 8)  # Using k=8 from Gap statistic method

# Display number of members in each cluster for both k = 7 and k = 8
print("Cluster distribution for k = 7:")
print(table(sil))

print("Cluster distribution for k = 8:")
print(table(gap))

```

```{r}
# Plotting the dendrogram with clusters indicated (k=8)
plot(ward, cex = 0.6)
rect.hclust(ward, k = 8, border = 2:8)  # Draw rectangle around each cluster

library(factoextra)

# Visualizing the clusters using fviz_cluster
fviz_cluster(list(data = df.scaling, cluster = sil))  # For k = 7
fviz_cluster(list(data = df.scaling, cluster = gap))  # For k = 8

# Plotting dendrogram with colors for the clusters (for k = 8)
ward.d <- as.dendrogram(ward)
ward.d <- color_branches(ward.d, k = 8)
ward.d <- color_labels(ward.d, k = 8)
plot(ward.d, main = "Cluster Dendrogram (Colored)", ylab = "Distance")
```
#Evaluation
```{r}
# First, add the hcluster labels to your dataframe
df$hclusters <- as.factor(gap)  # Use the clustering result from the hierarchical cut

# Summarizing the mean values for each cluster
hcluster_summary <- df %>%
  group_by(hclusters) %>%
  summarise(
    avg_age = mean(Age),
    avg_income = mean(Annual.Income..k..),
    avg_spending = mean(Spending.Score..1.100.)
  )


print(hcluster_summary)
```







##END















