---
title: "Q1"
author: "Nyew Jia Wen"
Student ID :210422078
output: html_document
---
### ===================== Q1a =====================
#set up a function
```{r}
f <- function(x){
  return(0.5*exp(-abs(x)))
}
```
#step 2
```{r}
MH <- function(N,s,initial_x){
  x <- numeric(N)#step 1 set up an initial value
  s>0
  x <- rep(s, N)
  x[1] <- initial_x
  for(i in 2:N){
    x_star <- rnorm(1,mean = x[i-1],sd = s)
    ratio <- f(x_star)/f(x[i-1])
    u <- runif(1,min=0,max=1)
    if (log(u)< log(ratio)){
      x[i] <- x_star
    }else{
        x[i] <- x[i-1]
      }
  }
  return(x)
}
```

```{r}
N <- 10000
s<-1
initial_x<-0
generated_samples <-MH(N,s ,initial_x)
```

#construct a histogram and kernel density plot
```{r}
hist(generated_samples,freq=FALSE,breaks = 30,col = "purple",main = "Metropolis-Hastings Sampling",xlab = "x",ylab = "Density")
curve(f(x),add = TRUE,col="cyan3",lwd=7)
kernel_dens <- density(generated_samples)
lines(kernel_dens$x, kernel_dens$y,col="darkorange",lwd=5)
legend("topright", legend = c("Histogram", "Kernel Density", "f(x)"), col = c("purple", "darkorange", "cyan3"), lty = c(1, 2, 1), lwd = 2)
```
#report the sample mean and standard deviation
```{r}
sample_mean<-mean(generated_samples)
sample_std<-sd(generated_samples)
cat("Sample Mean:",sample_mean,"\n")
cat("Sample Standard deviation:",sample_std,"\n")
```
### ===================== Q1b =====================

```{r}
compute_r_hat <- function(chains) {
  J <- length(chains)
  N <- length(chains[[1]])
  M <- sapply(chains, function(chain) mean(chain))
  V <- sapply(chains, function(chain) mean((chain - mean(chain))^2))
  W <- mean(V)
  M_bar <- mean(M)
  B <- mean((M - M_bar)^2)
  R_hat <- sqrt((B + W) / W)
  return(R_hat)
}
```

```{r}
N <- 2000
J <- 4
s_values <- seq(0.001, 1, by = 0.01)
R_hat_values <- numeric(length(s_values))
```

```{r}
for (i in seq_along(s_values)) {
  s <- s_values[i]
  chains <- lapply(1:J, function(j) {
    MH(N, s, initial_x)
  })
  R_hat_values[i] <- compute_r_hat(chains)
}
```

#provide a plot
```{r}
plot(s_values, R_hat_values, type = "l", xlab = "s", ylab = "R_hat", main = "R_hat vs. s")
```
---------------------------------------------------------------------------------------------------------
#-------------------------------------------------------Q2-----------------------------------------------

#data cleaning and reading
```{r}
library(DBI)#让R可以和SQLITE 说话
library(dplyr)#做数据清理的
#I used R with DBI and SQLite because the flight dataset was large, so it was more efficient to store the data in a database and query only the required fields.
library("RColorBrewer")
```

#load and write data to database
```{r}
if(file.exists("C:\\Users\\user\\Documents\\st2195\\airline.db"))#目的检查看有没有旧文档有的话删 怕影响新data
  file.remove("C:\\Users\\user\\Documents\\st2195\\airline.db")#后果：repeat data/different result
#总结：每次跑数据前都从干净数据库开始

conn <- dbConnect(RSQLite::SQLite(), "airline.db")
#conn就是数据库的把手
#I created a fresh SQLite database and established a connection so I could store and query large flight data efficiently without loading everything into memory at once. 
airports <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\airports.csv", header = TRUE)
planes <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\plane-data.csv", header = TRUE)


```

```{r}
#dbWriteTable（conn,name，value，expand的意思)
dbWriteTable(conn, "airports", airports,append=TRUE)
dbWriteTable(conn, "planes", planes,append=TRUE)
#总结：因为后面要用到sql去select和join

#因为数据太大 我们做partitioned loading分批读取
data.1998 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\1998.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)
data.1999 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\1999.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)
data.2000 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\2000.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)
data.2001 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\2001.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)
data.2002 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\2002.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)
data.2003 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\2003.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)
data.2004 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\2004.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)
data.2005 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\2005.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)
data.2006 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\2006.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)
data.2007 <- read.csv("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\2007.csv", header=TRUE, sep=',', stringsAsFactors = TRUE)

```

```{r}
data.1998$Cancelled <- factor(data.1998$Cancelled)
data.1999$Cancelled <- factor(data.1999$Cancelled)
data.2000$Cancelled <- factor(data.2000$Cancelled)
data.2001$Cancelled <- factor(data.2001$Cancelled)
data.2002$Cancelled <- factor(data.2002$Cancelled)
data.2003$Cancelled <- factor(data.2003$Cancelled)
data.2004$Cancelled <- factor(data.2004$Cancelled)
data.2005$Cancelled <- factor(data.2005$Cancelled)
data.2006$Cancelled <- factor(data.2006$Cancelled)
data.2007$Cancelled <- factor(data.2007$Cancelled)
#把他当factor来观察
#I treated cancellation as a categorical field because it represents flight status rather than a continuous numeric measure.
```

```{r}
dbWriteTable(conn, "flights", data.1998, append = TRUE)
dbWriteTable(conn, "flights", data.1999, append = TRUE)
dbWriteTable(conn, "flights", data.2000, append = TRUE)
dbWriteTable(conn, "flights", data.2001, append = TRUE)
dbWriteTable(conn, "flights", data.2002, append = TRUE)
dbWriteTable(conn, "flights", data.2003, append = TRUE)
dbWriteTable(conn, "flights", data.2004, append = TRUE)
dbWriteTable(conn, "flights", data.2005, append = TRUE)
dbWriteTable(conn, "flights", data.2006, append = TRUE)
dbWriteTable(conn, "flights", data.2007, append = TRUE)
```

### ===================== Q2a =====================
# Best time of day
```{r}
Time.LeastDelay <- dbGetQuery(conn,
  "SELECT DepTime
  FROM flights
  WHERE flights.Cancelled = 0 AND Diverted = 0 AND DepDelay > 0 #没取消的+没改降落地的+确实发生了depdelay
  ORDER BY DepDelay")#按小到大排序
#你排除了准点和提早的航班
head(Time.LeastDelay)
#总结：寻找在所有发生过 delay 的航班里，谁的 delay 最小”
#I filtered out cancelled and diverted flights, then examined departure times associated with low delay values and also compared frequencies of no-delay departures to identify time periods that were less delay-prone.
```

```{r}
Time.NoDelay <- dbGetQuery(conn,
  "SELECT DepTime, Count(*) AS total
  FROM flights
  WHERE flights.Cancelled = 0 AND Diverted = 0 AND DepDelay == 0
  GROUP BY DepTime#按出发时间分组
  ORDER BY total")#按数量从小到大排序
tail(Time.NoDelay)#取最后几行，也就是准点起飞次数最多的时间
#总结：找no delay最常见时间
#原因：
#1.一天才刚开始没被打乱flight has not been discrupted
#2.连锁延误还没累积 chain of delays x accumulated
#3.机场资源早高峰很稳定morning rush hour is stable
```

# Best day of the week
```{r}
Day.LeastDelay <- dbGetQuery(conn,
  "SELECT DayofWeek
  FROM flights
  WHERE flights.Cancelled = 0 AND Diverted = 0 AND DepDelay > 0
  ORDER BY DepDelay")
head(Day.LeastDelay)
#最小正延误对应的是星期几
```

```{r}
Day.NoDelay <- dbGetQuery(conn,
  "SELECT DayofWeek, Count(*) AS total
  FROM flights
  WHERE flights.Cancelled = 0 AND Diverted = 0 AND DepDelay == 0 
  GROUP BY DayofWeek
  ORDER BY total")
tail(Day.NoDelay)
#最不容易 delay 的日子
```

### ===================== Q2b =====================
#Do older planes suffer more delays?

# Average DepDelay
```{r}
AvgDepDelay <- dbGetQuery(conn,
 "SELECT planes.year AS planesyear, AVG(flights.DepDelay) AS AvgDelay
 FROM planes JOIN flights USING (tailnum)
 #你想研究“飞机年份”和“延误”的关系，必须把两张表连起来
 WHERE flights.Cancelled = 0 AND flights.Diverted = 0
 AND flights.DepDelay > 0
 #在已经延误的航班里，不同飞机年份的平均延误程度
 GROUP BY planesyear
 ORDER BY planesyear")
```

```{r}
AvgDepDelay$planesyear <- as.integer(AvgDepDelay$planesyear)
AvgDepDelay <- na.omit(AvgDepDelay)
AvgDepDelay <- AvgDepDelay[-1,]
```

```{r}
plot(AvgDepDelay$planesyear, AvgDepDelay$AvgDelay,
 pch = 4, xlab="Year Issued", ylab="Average Departure Delay in minutes",
 main="Flights with departure delay")
lines(AvgDepDelay$planesyear, AvgDepDelay$AvgDelay, col = "blue")#把点连成线，方便看趋势
abline(lm(AvgDepDelay$AvgDelay ~ AvgDepDelay$planesyear), col = "red")#把这条回归线画到图上
#如果红线是往下，表示：飞机年份越新平均 delay 越低也就是：越旧的飞机，delay 越高
#看1970和2000就知道旧机延误严重
```

# Average ArrDelay
```{r}
AvgArrDelay <- dbGetQuery(conn,
 "SELECT planes.year AS planesyear, AVG(flights.ArrDelay) AS AvgDelay
 FROM planes JOIN flights USING (tailnum)
 WHERE flights.Cancelled = 0 AND flights.Diverted = 0
 AND flights.ArrDelay > 0
 GROUP BY planesyear
 ORDER BY planesyear")
#旧飞机是否 arrival delay 也更严重
```

```{r}
AvgArrDelay$planesyear <- as.integer(AvgArrDelay$planesyear)
AvgArrDelay <- na.omit(AvgArrDelay)
AvgArrDelay <- AvgArrDelay[-1,]
```

```{r}
plot(AvgArrDelay$planesyear, AvgArrDelay$AvgDelay,
 pch = 4, xlab="Year Issued", ylab="Average Arrival Delay in minutes",
 main="Flights with arrival delay")
lines(AvgArrDelay$planesyear, AvgArrDelay$AvgDelay, col = "blue")
abline(lm(AvgArrDelay$AvgDelay ~ AvgArrDelay$planesyear), col = "red")
#发现新机抵达延迟，起飞时大家新旧无分别（红线往上了一点）
#怀疑理由：被派去更远的地方·，更挤的航线，所以落地时候排队堵车

```

### ===================== Q2c =====================
```{r}
library(ggplot2)
```
# Define a function to fit the model and get coefficients for a given year
# Fit the logistic regression model
```{r}
#为什么要比较年份间的系数变化因为它想知道：同一个变量，在不同年份的影响是不是一样？

#专门写一个函数处理某一年的数据
#为什么要这样？因为我要对不同年份重复做同一件事
get_coefficients <- function(year) {
  df_year <- read.csv(paste0("C:\\Users\\user\\Documents\\st2195\\EXTRACT\\", year, ".csv"))#为什么要检查diverted有没有两种价值呢01是正常的 0是不正常的模型没法学习
  if (length(unique(df_year$Diverted)) < 2) {
    print(paste("Data for year", year, "contains only one class in 'Diverted'. Skipping..."))
    return(NULL)
  }
  model <- glm(Diverted ~ DepTime + CRSDepTime + CRSArrTime + Distance + UniqueCarrier, data = df_year, family = binomial)
  return(data.frame(Year = year, Coefficients = coef(model)))#因为你要把每一年的模型系数整理成统一格式，后面才能合并
  #coeff正就增加DepTime = 0.002起飞时间越往后，diverted 概率可能略微增加
  #coeff负CarrierAA = -0.5相比基准航空公司，AA 的 diverted 倾向更低
}
  
  
  

```

# Apply the function to each chosen year
```{r}
coefficients_list <- lapply(c(2001,2004,2006), get_coefficients)#对 2001、2004、2006 这几个年份，一个一个去跑 get_coefficients
coefficients <- do.call(rbind, coefficients_list)
print(coefficients)
#In my implementation, I focused on selected years that had both diverted and non-diverted cases so that the logistic model could be estimated properly.
```

# Convert the coefficients data frame to a format suitable for ggplot
```{r}
coefficients_long <- tidyr::gather(coefficients, key = "Variable", value = "Coefficient", -Year)#把数据转成长格式，方便 ggplot 作图

ggplot(coefficients_long, aes(x = Year, y = Coefficient, color = Variable)) +
  geom_line() +
  labs(title = "Logistic Regression Coefficients Over Time",
       x = "Year",
       y = "Coefficient",
       color = "Variable")
```







```
```















